The Talking Heads Experiment

### 2015-04-23 chap2/chap2.tex
edited todo notes

Agents can only engage in interactions with other agents 
when they are physically instantiated
in a robot. Each agent has a basic brain architecture
with different layers performing the cognitive functions 
relevant for playing language games: 
\begin{itemize}
\item A perceptual layer which performs low-level signal 
processing to segment the image and collect data about 
each segment such as the colour, size, position or shape
of a segment. 
\item A conceptual layer which categorises and 
conceptualises the segmented and processed image. 
It is based on a self-generated and evolving repertoire
of categorial distinctions, such as red versus green, or
small versus large. Such a repertoire is referred to as 
the agent's ontology in this book. 
\item A lexical layer which maintains an evolving repertoire
of associations between meanings and words, which I will refer to 
as the lexicon, and performs lexical lookup while parsing 
or producing utterances. 
\item A syntactic layer which uses 
grammatical schemata for organising words
in larger structures or for recognising these structures and
reconstructing complex meanings. 
\item A pragmatic layer which carries out the
scripts for playing language games and maintains the machinery 
for engaging in interactions with other agents in a 
shared environment. 
\end{itemize}
Each of these layers is described in more detail 
later. The internals of the layers are not static but
constantly evolving and adapting. They are not strictly 
modular but coupled in various ways to each other.
Each verbal interaction 
in effect changes the agent's internal state and thus 
influences future behaviour. A new, virgin agent 
starts without any built-in ontology, lexicon, or grammar. 
This is one of the crucial points of the whole experiment, 
because we want to test possible theories on how language
and meaning can evolve and be acquired ab initio\todo{GSR: no italics in commonly used loan words}.

@@

The environment of the Talking Heads is 
not fixed. The human experimenter 
may change the position of objects, add new kinds of 
objects, or eliminate
others. Consequently a strategy of naming individual objects
will not work. It would lead to a proliferation of 
proper names and it would require the Talking Heads to
recognise objects, which is very difficult to do.\footnote{
For a thorough exposition of the difficulties of 
object recognition, see \cite{Ullman:1996}. 
Object constancy comes fairly late in the 
acquisition of a child's ontology, as Piaget's conservation
experiments have shown. Language probably plays an important
role in forming the notion of an object.}
Indeed, humans don't exclusively use proper names in 
natural language conversations 
either. We say `could you give me the red small square' as 
opposed to `could you give me O\_143'. Natural language \todo{inconsistency: we have 'O\_143' here but 'O143' on the next page}
words like `red' or `small' name perceptually grounded categories
and syntactic structures indicate how they should
be combined and used to find the topic. The relation between 
a language expression and its referent is therefore 
always indirect. This is summarised
in the semiotic square (\figref{triangle}), which I will use throughout this book to 
help understand and analyse the nature of language communication. \is{semiotic square}
The semiotic square relates the four entities
involved in a verbal interaction: 
\begin{itemize}
\item An {\itshape utterance}, such as `small red square', which is 
transmitted as a physical 
signal from one agent to another one through the 
external environment. It is written between double quotes. 
\item A {\itshape meaning}, which consists of categories like [\textsc{red}], [\textsc{small}], 
or combinations of categories, like \{[\textsc{red}] [\textsc{small}]\}. Labels of 
categories are written in capital letters between 
square brackets. 
\item An {\itshape image segment}, denoted as \emph{S143}, which is 
a segment of the image perceived through the camera.
\item A {\itshape referent}, like object \emph{O143}, which is 
an entity in the real world. 
\end{itemize}

>>>>

Agents can only engage in interactions with other agents 
when they are physically instantiated
in a robot. Each agent has a basic brain architecture
with different layers performing the cognitive functions 
relevant for playing language games: 
\begin{itemize}
\item A perceptual layer which performs low-level signal 
processing to segment the image and collect data about 
each segment such as the colour, size, position or shape
of a segment. 
\item A conceptual layer which categorises and 
conceptualises the segmented and processed image. 
It is based on a self-generated and evolving repertoire
of categorial distinctions, such as red versus green, or
small versus large. Such a repertoire is referred to as 
the agent's ontology in this book. 
\item A lexical layer which maintains an evolving repertoire
of associations between meanings and words, which I will refer to 
as the lexicon, and performs lexical lookup while parsing 
or producing utterances. 
\item A syntactic layer which uses 
grammatical schemata for organising words
in larger structures or for recognising these structures and
reconstructing complex meanings. 
\item A pragmatic layer which carries out the
scripts for playing language games and maintains the machinery 
for engaging in interactions with other agents in a 
shared environment. 
\end{itemize}
Each of these layers is described in more detail 
later. The internals of the layers are not static but
constantly evolving and adapting. They are not strictly 
modular but coupled in various ways to each other.
Each verbal interaction 
in effect changes the agent's internal state and thus 
influences future behaviour. A new, virgin agent 
starts without any built-in ontology, lexicon, or grammar. 
This is one of the crucial points of the whole experiment, 
because we want to test possible theories on how language
and meaning can evolve and be acquired ab initio\todo{removed italics here}.

@@

The environment of the Talking Heads is 
not fixed. The human experimenter 
may change the position of objects, add new kinds of 
objects, or eliminate
others. Consequently a strategy of naming individual objects
will not work. It would lead to a proliferation of 
proper names and it would require the Talking Heads to
recognise objects, which is very difficult to do.\footnote{
For a thorough exposition of the difficulties of 
object recognition, see \cite{Ullman:1996}. 
Object constancy comes fairly late in the 
acquisition of a child's ontology, as Piaget's conservation
experiments have shown. Language probably plays an important
role in forming the notion of an object.}
Indeed, humans don't exclusively use proper names in 
natural language conversations 
either. We say `could you give me the red small square' as 
opposed to `could you give me O\_143'. Natural language
words like `red' or `small' name perceptually grounded categories
and syntactic structures indicate how they should
be combined and used to find the topic. The relation between 
a language expression and its referent is therefore 
always indirect. This is summarised
in the semiotic square (\figref{triangle}), which I will use throughout this book to 
help understand and analyse the nature of language communication. \is{semiotic square}
The semiotic square relates the four entities
involved in a verbal interaction: 
\begin{itemize}
\item An {\itshape utterance}, such as `small red square', which is 
transmitted as a physical 
signal from one agent to another one through the 
external environment. It is written between double quotes. 
\item A {\itshape meaning}, which consists of categories like [\textsc{red}], [\textsc{small}], 
or combinations of categories, like \{[\textsc{red}] [\textsc{small}]\}. Labels of 
categories are written in capital letters between 
square brackets. 
\item An {\itshape image segment}, denoted as \emph{S143}, which is 
a segment of the image perceived through the camera.
\item A {\itshape referent}, like object \emph{O143}, which is 
an entity in the real world. 
\end{itemize}

### 2015-04-23 chap4/chap4.tex
edited todo notes

The \todo{This intro is way to long, please include another \textbackslash section} previous chapter laid the perceptual groundwork 
for investigating conceptualisation and language 
communication. It proposed a set of mechanisms for segmenting a 
scene and extracting characteristics about each 
segment. The next task of the agent is to use these results
from the perceptual layer to categorise 
and hence conceptualise the
scene. For the time being, I will 
ignore the full complexity of natural
language meaning and focus only on the most 
simple type of conceptualisation one can
imagine, namely categories, which logicians refer 
to as unary predicates, as well as conjunctive 
combinations of categories. Words like `blue', `light' or
`square' name such categories. We not only need a way 
that the Talking Heads can use such categories to conceptualise
a scene based on visual perception, but we also need 
mechanisms that can explain how such categories might 
develop or be acquired by each agent autonomously and 
without supervised training. This is clearly an enormous
challenge and no known universally accepted solution 
exists to this problem.

>>>>

The previous chapter laid the perceptual groundwork 
for investigating conceptualisation and language 
communication. It proposed a set of mechanisms for segmenting a 
scene and extracting characteristics about each 
segment. The next task of the agent is to use these results
from the perceptual layer to categorise 
and hence conceptualise the
scene. For the time being, I will 
ignore the full complexity of natural
language meaning and focus only on the most 
simple type of conceptualisation one can
imagine, namely categories, which logicians refer 
to as unary predicates, as well as conjunctive 
combinations of categories. Words like `blue', `light' or
`square' name such categories. We not only need a way 
that the Talking Heads can use such categories to conceptualise
a scene based on visual perception, but we also need 
mechanisms that can explain how such categories might 
develop or be acquired by each agent autonomously and 
without supervised training. This is clearly an enormous
challenge and no known universally accepted solution 
exists to this problem.

### 2015-04-23 chap9/chap9.tex
edited todo notes

N01SE celebrates the essential excess from which information is drawn. It probes many different ways of seeing and 
being in the world. Chances are your own sense of order is already someone elses N01SE.\todo{quotations are not set in italics, see GSR}
\end{quotation}
%\begin{figure}[htbp]
%  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/latour.pdf}}
%\caption{\label{latour} 
%Luc Steels Discusses The N01SE exhibition with Bruno Latour who intensively followed the experiments already from the first series at the Laboratorium exhibition in Antwerp. His comments significantly improved the initial designs.}
%\end{figure}

@@

\todo{removed one \textbackslash rule from the last entry. Also, why is \textbackslash rule set to .8 of textwidth?!}
\enlargethispage{1\baselineskip}

>>>>

N01SE celebrates the essential excess from which information is drawn. It probes many different ways of seeing and 
being in the world. Chances are your own sense of order is already someone elses N01SE.\todo{I have unset italics for this quotation}
\end{quotation}
%\begin{figure}[htbp]
%  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/latour.pdf}}
%\caption{\label{latour} 
%Luc Steels Discusses The N01SE exhibition with Bruno Latour who intensively followed the experiments already from the first series at the Laboratorium exhibition in Antwerp. His comments significantly improved the initial designs.}
%\end{figure}

@@

\enlargethispage{1\baselineskip}
